---
title: "p8105_hw1_geg2145"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(moderndive)
```

# Problem 1

```{r loading_data}
data("early_january_weather")
weather = early_january_weather
```

This dataset has hourly meterological data for the month of January 2013 for three airports: La Guardia, John F. Kennedy, and Newark Liberty. In total, this dataset has `r nrow(weather)` rows and `r ncol(weather)` columns. The `origin` variable tells us the weather station that made the recording, although for this subset, all recordings are made from "EWR" (Newark). We also have the recorded `temperature` and `relative humidity`, along with the `time` the recording was made (date and hour). The mean temperature across all the recordings available is `r round(mean(pull(weather,temp)), 3)` degrees Fahrenheit.

```{r}
ggplot(weather) + 
  geom_point(aes(x= time_hour, y = temp, col= humid)) + 
  labs(x="Time", y="Temperature (in F)", 
       col="Relative Humidity", 
       title="January got warmer and more humid as time went on")

ggsave("problem1_plot.png")
```

In the above scatterplot, we see `time` in the x-axis and `temperature` in the y axis. Although the temperature fluctuates throughout the day (as we go from day to night, and back to day), we see that overall, temperature is trending upwards as time goes on. We also see our points colored by their `relative humidity`, and see most dark points in earlier dates, with most of the lighter ones in later dates. This indicates that generally, the days got more humid as January went along.

